Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-10 paper.qmd
Updated the code

```{r}
#| label: load-libraries
#| warning: false
#| echo: false
# remotes::install_github("JayaniLakshika/quollr")
library(quollr)
library(tibble)
library(dplyr)
# remotes::install_github("jlmelville/snedata")
# library(snedata)
# library(ggflowchart)
#library(purrr) ## map function
#library(gridExtra) ## for grid.arrange
# library(rsample)
# library(DT)
#library(ggbeeswarm)
library(ggplot2)
library(readr)
library(tidyr)

# library(Rtsne)
# library(uwot)
# library(phateR)
library(patchwork)
library(png)
#library(langevitour)
library(colorspace)
library(kableExtra)
#library(grid)
library(conflicted)
conflicts_prefer(dplyr::filter)
```

```{r}
#| label: plot-theme
theme_set(theme_linedraw() +
   theme(
     #aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
     panel.background = element_rect(fill = 'transparent', 
                                     colour = NA),
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(), 
     axis.title.x = element_blank(), axis.title.y = element_blank(),
     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
     legend.background = element_rect(fill = 'transparent', 
                                      colour = NA),
     legend.key = element_rect(fill = 'transparent', 
                               colour = NA),
     legend.position = "bottom", 
     legend.title = element_blank(), 
     legend.text = element_text(size=4),
     legend.key.height = unit(0.25, 'cm'),
     legend.key.width = unit(0.25, 'cm')
   )
)
interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
  annotation_custom(grid::textGrob(label = label,
      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
      gp = grid::gpar(cex = cex, col=col)))
}
```

@@

sc_ltr_pos <- c(0.08, 0.9)
# sc_xlims <- c(-0.5, 1.47)
# sc_ylims <- c(-0.32, 2.1)
sc_xlims <- c(-0.3, 1.2)
sc_ylims <- c(-0.17, 1.25)

nldr_two_curvy <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos, cex = 1.5)

nldr_two_curvy2 <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
  interior_annotation("a", sc_ltr_pos, cex = 1.5)

nldr_two_curvy_original <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
  theme(aspect.ratio = 1) +
  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
  interior_annotation("a", sc_ltr_pos)

nldr_two_curvy_original2 <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
  theme(aspect.ratio = 1) +
  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
  interior_annotation("a1", sc_ltr_pos)
```

>>>>

```{r}
#| label: load-libraries
#| warning: false
#| echo: false
library(quollr)
library(tibble)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(patchwork)
library(png)
library(colorspace)
library(kableExtra)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
```

```{r}
#| label: plot-theme
theme_set(theme_linedraw() +
   theme(
     #aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
     panel.background = element_rect(fill = 'transparent', 
                                     colour = NA),
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(), 
     axis.title.x = element_blank(), axis.title.y = element_blank(),
     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
     legend.background = element_rect(fill = 'transparent', 
                                      colour = NA),
     legend.key = element_rect(fill = 'transparent', 
                               colour = NA),
     legend.position = "bottom", 
     legend.title = element_blank(), 
     legend.text = element_text(size=4),
     legend.key.height = unit(0.25, 'cm'),
     legend.key.width = unit(0.25, 'cm')
   )
)

interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
  annotation_custom(grid::textGrob(label = label,
      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
      gp = grid::gpar(cex = cex, col=col)))
}
```

@@

sc_ltr_pos <- c(0.08, 0.9)
sc_xlims <- c(-0.3, 1.2)
sc_ylims <- c(-0.17, 1.25)

nldr_two_curvy_original <- ggplot(tsne_two_curvy_scaled, aes(x = tSNE1, y = tSNE2)) +
  geom_point(alpha = 0.5, color = "#636363", size = 0.5)

# First plot with xlim and ylim
nldr_two_curvy <- nldr_two_curvy_original+
  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
  xlim(sc_xlims) +
  ylim(sc_ylims)

# Second plot (identical to base except for xlim/ylim)
nldr_two_curvy2 <- nldr_two_curvy_original +
  interior_annotation("a", sc_ltr_pos, cex = 1.5)

# Third plot with aspect ratio and different annotation
nldr_two_curvy_original2 <- nldr_two_curvy_original +
  theme(aspect.ratio = 1) +
  interior_annotation("a1", sc_ltr_pos)
```
